Slight Positive Correlation Between Gendered and Racialized Police Strip Searches in the Toronto Municipal, 2020-2021

### 2024-01-25 outputs/paper/paper.qmd
update results

The other dataset conveys all police arrests that involve strip searching and the reasons behind strip searching. Under close examination, the naming of individual data cells in different columns from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`. Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on. After renaming, the big dataset was split into containing relevant columns ([see @tbl-cleaned_search_reasons_data]). Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values for all the `reason` column.

@@

Furthermore, the reason for keeping rows of events that don't include strip search is to later produce summary statistics of events with and without strip searches. This further aligns with the task of the paper at hand, which is to critically reflect on the rate of strip searching on people with racialized and gendered identities.

@@

@@

```{r fig.height=6, fig.width=7}
#| echo: false
#| message: false
#| label: fig-race_strip_search
#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021

  cleaned_race_gender_data |>
  group_by(race, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  ggplot(aes(x = race, y = number_of_arrests, fill = strip_search)) + 
  geom_bar(stat="identity") +
  theme(legend.key.size = unit(0.5, 'cm'), axis.text.x = element_text(angle = 70, vjust = 1, hjust= 1)) +
  labs(x = "Race",
       y = "Number of Arrests",
       fill = "Strip Search") + scale_fill_viridis_d(option="cividis")

@@

cleaned_race_gender_data |>
  group_by(race, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  filter(strip_search == "Yes") |> 
  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
  geom_bar() + 
  labs(y = "Number of Arrests", x = "Strip Search", fill = "Strip Search") +
  facet_wrap(~race) + scale_fill_viridis_d(option="cividis")

@@

  cleaned_race_gender_data |>
  group_by(gender, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  ggplot(aes(x = gender, y = number_of_arrests, fill = strip_search)) + 
  geom_bar(stat="identity") + 
  coord_flip() +
  theme(legend.key.size = unit(0.5, 'cm')) +
  labs(x = "Gender",
       y = "Number of Arrests",
       fill = "Strip Search"
  ) + scale_fill_viridis_d(option="cividis")

@@

#create a scatterplot to show each observation in the dataset, comparing the total reasons to racial identitiy found in each strip search
#colour the points based on the perceived race of the captive
#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
cleaned_search_reasons_data |>
  group_by(total_reasons, race) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = race)) +
  geom_line(alpha = 0.5) +
  geom_point(aes(size = count), alpha = 0.5) +
  labs(x = "Total Reasons for Strip Searching",
       y = "Count",
       color = "Race"
       ) +
  theme(legend.key.size = unit(0.5, 'cm')) +
  scale_fill_viridis_d(option="cividis")

@@

cleaned_search_reasons_data |>
  group_by(total_reasons, gender) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = gender)) +
  geom_line(alpha = 0.5) +
  geom_point(aes(size = count), alpha = 0.5) +
  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
  theme(legend.key.size = unit(0.5, 'cm')) +
  scale_fill_viridis_d(option="cividis")

@@

>>>>

To further inspect the nature of each strip search, reasons for strip searching prove a necessary data source. Thus, this other dataset conveys all police arrests that involve strip searching and the reasons behind strip searching. Under close examination, the naming of individual data cells in different columns from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA`. Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on. After renaming, the big dataset was split into containing relevant columns ([see @tbl-cleaned_search_reasons_data]). Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for rows with `NA` as values for all the `reason` column.

@@

Throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for 12% of all arrests [see @tbl-reasons-strip-search]. Excluding those whose racial identities are unknown and non-existent, by percentage, the three most affected racial groups are, in descending order, of Indigenous, African and Caucasian descent, standing at 16%, 14% and 13% each [see @tbl-race-strip-search]. Gender-wise, men are more likely to be strip searched compared to women, at 12% and 10% respectively (@tbl-gender-strip-search). Though, observably, the difference between the two sexes quite marginal. People of unidentified gender identity aren't strip searched at all according to the dataset (denoted as `U` in the tables and graphs).

As of January 22, 2023, there are 434 COVID-19 immunization clinics across the City of Toronto. Further breakdown shows that there are 5 clinics that are city operated, 16 clinics that are run by hospitals, and 415 are offered in local pharmacies. On average, there are 17 clinics per ward in Toronto, with standard deviation of 6. The wards with the highest number of clinics are: Toronto Center (Ward 13) with 29 clinics, Etobicoke-Lakeshore (Ward 3) with 28 clinics, and University-Rosedale (Ward 11) with 28 clinics. The wards with the lowest number of clinics are: Humber River-Black Creek (Ward 7) with 9 clinics, Davenport (Ward 9) with 12 clinics, Scarborough Southwest (Ward 20) with 12 clinics, Scarborough-Agincourt (Ward 22) with 12 clinics, and Scarborough-Rouge Park (Ward 25) with 12 clinics.

In add

@@

In addition, a new column called "Total Reasons" totals all reasons used to account for each police strip search is added with `mutate()`, where all instances of `1` and `0` are changed into numeric with `as.numeric()` and then summed together.
to ensure that all data are comparative to one another

@@

```{r fig.width=8, fig.height=5}
#| echo: false
#| message: false
#| label: fig-race_strip_search
#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021

cleaned_race_gender_data |>
  group_by(race, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  group_by(race) |>
  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
  ggplot(aes(x = race, y = percentage, fill = strip_search)) + 
  geom_bar(stat="identity") +
  coord_flip() +
  theme_minimal() +
  theme(legend.key.size = unit(0.5, 'cm')) +
  labs(x = "Race",
       y = "Percentage",
       fill = "Strip Search") + scale_fill_viridis_d(option="cividis")

@@

cleaned_race_gender_data |>
  group_by(race, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  group_by(race) |>
  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
  ggplot(mapping = aes(x = gender, fill = strip_search), data=cleaned_race_gender_data) +
  geom_bar() + 
  labs(y = "Percentage", x = "Strip Search", fill = "Strip Search") +
  facet_wrap(~race) + scale_fill_viridis_d(option="cividis") +

@@

  cleaned_race_gender_data |>
  group_by(gender, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  ggplot(aes(x = gender, y = number_of_arrests, fill = strip_search)) + 
  geom_bar(stat="identity") + 
  coord_flip() +
  theme_minimal() +
  theme(legend.key.size = unit(0.5, 'cm')) +
  labs(x = "Gender",
       y = "Number of Arrests",
       fill = "Strip Search"
  ) + scale_fill_viridis_d(option="cividis")

cleaned_race_gender_data |>
  group_by(gender, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  group_by(gender) |>
  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
  ggplot(aes(x = gender, y = percentage, fill = strip_search)) + 
  geom_bar(stat="identity") + 
  coord_flip() +
  theme_minimal() +
  theme(legend.key.size = unit(0.5, 'cm')) +
  labs(x = "Gender",
       y = "Percentage",
       fill = "Strip Search"
  ) + scale_fill_viridis_d(option="cividis")

cleaned_race_gender_data |>
  group_by(gender, strip_search) |>
  count() |>
  rename(number_of_arrests = n) |>
  group_by(gender) |>
  mutate(percentage = round(number_of_arrests/sum(number_of_arrests)*100)) |>
  ggplot(mapping = aes(x = strip_search, fill = strip_search), data=cleaned_race_gender_data) +
  geom_bar() + 
  labs(y = "Number of Arressts", x = "Strip Search", fill = "Strip Search") +
  facet_wrap(~gender) + scale_fill_viridis_d(option="cividis")

@@

#create a scatterplot to show each observation in the dataset, comparing the total reasons to racial identitiy found in each strip search
#colour the points based on the perceived race of the captive
#code for geom_point() referenced from https://github.com/TDonofrio62/Shooting-Occurrences-in-Toronto/blob/main/Shooting_Occurrences/outputs/paper/Shooting_Occurrences.Rmd
cleaned_search_reasons_data |>
  group_by(total_reasons, race) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = race)) +
  geom_line(alpha = 0.5) +
  geom_point(aes(size = count), alpha = 0.5) +
  labs(x = "Total Reasons for Strip Searching",
       y = "Count",
       color = "Race"
       ) +
  theme_minimal() +
  theme(legend.key.size = unit(0.5, 'cm')) +
  scale_fill_viridis_d(option="cividis")

@@

cleaned_search_reasons_data |>
  group_by(total_reasons, gender) |>
  count() |>
  rename(count = n) |>
  ggplot(aes(x = total_reasons, y = count, colour = gender)) +
  geom_line(alpha = 0.5) +
  geom_point(aes(size = count), alpha = 0.5) +
  labs(y = "Count", x = "Total Reasons for Strip Searching", color = "Gender") +
  theme(legend.key.size = unit(0.5, 'cm')) +
  theme_minimal() +
  scale_fill_viridis_d(option="cividis")

@@

binaristic view of gender
